---
title: "NorScir: data processing (analysebok)"
author: "Are Edvardsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The _NorScir_ package provide interactive results from the registry
_NorScir_ at _Rapporteket_.

```{r}
library(NorScir)
```

This package depend on the the _rapbase_ package. Underlying processing of 
registry data is describe in this document.

## Sample data
This package comes with a sample data set for testing purposes

```{r}
data(NorScirSampleData)
head(names(NorScirSampleData), 10)
```

It contains realistic data but with nonsense personal info. Sample data
should only be used during package development and never in a full scale
testing or production environment.


## General functions
These scripts contain supporting functions to be used by the report functions.

### NSLibUtvalg.R
Provided a (raw) set of data this function returns a subset based on the
criteria provided by its arguments 

### NSLoadRegData.R
Defines the SQL query and loads the global registry data set. Since this
function assumes real data it
will only run successfully in a test or production environment, or on a replica
of it.

In order to query registry data a unike registry identifier must be provided.
For NorScir registrey this idetifier is "NorScir" and must be assigned to the
variable _registyName_ withint the function _NSLoadRegData()_:

```{r, eval=FALSE}
registryName <- "NorScir"
```


## Report functions
Each _R_-function within the scripts listed below provide at least one, but
usually several, set of results at _Rapporteket_.

### NSFigAndelStabel.R
Provides figure of stacked bars, _e.g._:

```{r, fig.width=7}
FigAndelStabel(RegData=NorScirSampleData, outfile='', libkat='',
               valgtVar='NevrNivaaInnUt', datoFra='2010-01-01',
               datoTil='3000-01-01', minald=0, maxald=120, erMann='',
               traume='', enhetsUtvalg='0', reshID=12345, hentData=0)
```


### NSFigFordeling.R
Provide distribution as bars, _e.g._:
```{r, fig.width=7}
FigFordeling(RegData=NorScirSampleData, libkat='', outfile='', valgtVar='AAis',
             datoFra='2010-01-01', datoTil='2050-01-01', AIS='', minald=0,
             maxald=120, erMann='', traume='', enhetsUtvalg=0, reshID=12345,
             sml=1, hentData=0)
```

#### Analysis
For variables "AAis" and "FAis" an additional cathegory "None" was introduced
at some point. In the above function "None" is recoded as "U" as requested by
the registry during autumn 2015.


### NSFigMeanMed.R
Provide mean and median per hospital, _e.g._:
```{r, fig.width=7}
FigMeanMed(RegData=NorScirSampleData, valgtVar='Alder', valgtMaal='Gjsn',
           datoFra='2010-01-01', datoTil='2015-12-31', AIS='', minald=0,
           maxald=130, erMann='', traume='', libkat='', outfile='', hentData=0)
```


